Online Resources 1: Supplementary results and tables

### 2025-06-20 analysis/supplementary-materials/ESM_1.qmd
Article review. Statistical analysis and stats changed to separate flakes from elongated blanks.

>>>>

```{r}
#| label: stats-wilcox-review
#| include: false

# Terrace wilcox statistics
## DB prep
wilcox_t <- phase_dataset |> 
  dplyr::mutate(Prov = case_when(
    RMTYPE == "T1" | RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "Local",
    RMTYPE == "T6" | RMTYPE == "T7" | RMTYPE == "T8" | RMTYPE == "T9" ~ "Non-local",
    TRUE ~ RMTYPE)) |> 
  filter(Prov %in% c("Local", "Non-local")) |> 
  select(ID, CLASS_ELONG, THICKNESS, MAXWIDTH, LENGTH, WEIGHT, Prov, Level)

# Proto-Solutrean flakes statistical test
## NL cores were not represented enough (n) for statistics
wt_flake_proto <- wilcox_t |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  filter(CLASS_ELONG == "Flake")

## Flake weight (fw)
dplyr::group_by(wt_flake_proto, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_fw_proto <- wilcox.test(WEIGHT ~ Prov, data = wt_flake_proto, exact = FALSE)
## Flake length (fl)
dplyr::group_by(wt_flake_proto, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_fl_proto <- wilcox.test(LENGTH ~ Prov, data = wt_flake_proto, exact = FALSE)
## Flake width (fwi)
dplyr::group_by(wt_flake_proto, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_fwi_proto <- wilcox.test(MAXWIDTH ~ Prov, data = wt_flake_proto, exact = FALSE)

# Proto-Solutrean elongated blanks statistical test
wt_elong_proto <- wilcox_t |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  filter(CLASS_ELONG == "ElongatedProd")

## Elongated blank weight (ew)
dplyr::group_by(wt_elong_proto, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_ew_proto <- wilcox.test(WEIGHT ~ Prov, data = wt_elong_proto, exact = FALSE)
## Elongated blank length (el)
dplyr::group_by(wt_elong_proto, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_el_proto <- wilcox.test(LENGTH ~ Prov, data = wt_elong_proto, exact = FALSE)
## Elongated blank width (ewi)
dplyr::group_by(wt_elong_proto, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_ewi_proto <- wilcox.test(MAXWIDTH ~ Prov, data = wt_elong_proto, exact = FALSE)

# Gravettian flakes statistical test
wt_flake_grav <- wilcox_t |> 
  filter(Level %in% c("6", "7")) |> 
  filter(CLASS_ELONG == "Flake")

## Flake weight
dplyr::group_by(wt_flake_grav, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_fw_grav <- wilcox.test(WEIGHT ~ Prov, data = wt_flake_grav, exact = FALSE)
## Flake length
dplyr::group_by(wt_flake_grav, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_fl_grav <- wilcox.test(LENGTH ~ Prov, data = wt_flake_grav, exact = FALSE)
## Flake width
dplyr::group_by(wt_flake_grav, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_fwi_grav <- wilcox.test(MAXWIDTH ~ Prov, data = wt_flake_grav, exact = FALSE)

# Gravettian elongated blanks statistical test
wt_elong_grav <- wilcox_t |> 
  filter(Level %in% c("6", "7")) |> 
  filter(CLASS_ELONG == "ElongatedProd")

## Elongated blanks weight
dplyr::group_by(wt_elong_grav, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_ew_grav <- wilcox.test(WEIGHT ~ Prov, data = wt_elong_grav, exact = FALSE)
## Elongated blanks length
dplyr::group_by(wt_elong_grav, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_el_grav <- wilcox.test(LENGTH ~ Prov, data = wt_elong_grav, exact = FALSE)
## Elongated blanks width
dplyr::group_by(wt_elong_grav, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_ewi_grav <- wilcox.test(MAXWIDTH ~ Prov, data = wt_elong_grav, exact = FALSE)

# Shelter wilcox statistics
## DB prep
wilcox_t_shelter <- solutrean_dataset |> 
  dplyr::mutate(Prov = case_when(
    RMTYPE == "T1" | RMTYPE == "T2" | RMTYPE == "T3" | RMTYPE == "T4" | RMTYPE == "T5" ~ "Local",
    RMTYPE == "T6" | RMTYPE == "T7" | RMTYPE == "T8" | RMTYPE == "T9" | RMTYPE == "T11" ~ "Non-local",
    TRUE ~ RMTYPE)) |> 
   mutate(CLASS = case_when(
    CLASS == "ElongBlank" ~ "ElongatedProd",
    TRUE ~ CLASS)) |>
  filter(Prov %in% c("Local", "Non-local")) |> 
  select(ID, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT, Prov, LEVEL)

# Solutrean flakes statistical test
wt_flake_s <- wilcox_t_shelter |>
  filter(CLASS == "Flake")

## Flake weight
dplyr::group_by(wt_flake_s, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_fw_s <- wilcox.test(WEIGHT ~ Prov, data = wt_flake_s, exact = FALSE)
## Flake length
dplyr::group_by(wt_flake_s, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_fl_s <- wilcox.test(LENGTH ~ Prov, data = wt_flake_s, exact = FALSE)
## Flake width
dplyr::group_by(wt_flake_s, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_fwi_s <- wilcox.test(MAXWIDTH ~ Prov, data = wt_flake_s, exact = FALSE)

# Solutrean elongated blanks statistical test
wt_elong_s <- wilcox_t_shelter |>
  filter(CLASS == "ElongatedProd")

## Elongated blanks weight
dplyr::group_by(wt_elong_s, Prov) |> 
  summarise(count = n(), median = median(WEIGHT, na.rm = TRUE))
res_ew_s <- wilcox.test(WEIGHT ~ Prov, data = wt_elong_s, exact = FALSE)
## Elongated blanks length
dplyr::group_by(wt_elong_s, Prov) |> 
  summarise(count = n(), median = median(LENGTH, na.rm = TRUE))
res_el_s <- wilcox.test(LENGTH ~ Prov, data = wt_elong_s, exact = FALSE)
## Elongated blanks width
dplyr::group_by(wt_elong_s, Prov) |> 
  summarise(count = n(), median = median(MAXWIDTH, na.rm = TRUE))
res_ewi_s <- wilcox.test(MAXWIDTH ~ Prov, data = wt_elong_s, exact = FALSE)

# Flake stats table
p.value_flake <- tibble(
  Technocomplex = c("Gravettian", " "," ", " ", "Proto-Solutrean", " ", " ", " ", "Solutrean", " ", " ", " "),
  Variable = c("Weight", "Length", "Width", "(n)","Weight", "Length", "Width","(n)", "Weight", "Length", "Width", "(n)"),
  L_median = c("3.2", "25.3", "21.4", "65", "2.5", "22.1", "18.8", "123", "1.9", "24", "18.2", "240"),
  NL_median = c("2", "20.9", "19.3", "60", "2.2", "19.1", "17.7", "36", "1.55", "20.7", "20.1", "65"),
  P_Value = round(c(res_fw_grav$p.value, res_fl_grav$p.value, res_fwi_grav$p.value, 0, res_fw_proto$p.value, res_fl_proto$p.value, res_fwi_proto$p.value, 0, res_fw_s$p.value, res_fl_s$p.value, res_fwi_s$p.value, 0), 5)) |>   mutate(Significance = case_when(
      P_Value <= 0.001 ~ "***",
      P_Value <= 0.01  ~ "**",
      P_Value <= 0.05  ~ "*",
      TRUE             ~ ""),
    P_value = paste0(format(P_Value, nsmall = 4), Significance)  # Keep decimal places
  ) |> 
  select(-P_Value, -Significance) |> 
  dplyr::mutate(P_value = case_when(
    P_value == "0.00000***" ~ "-",
    TRUE ~ P_value))

# Elongated blanks stats table
p.value_elong <- tibble(
  Technocomplex = c("Gravettian", " "," ", " ", "Proto-Solutrean", " ", " ", " ", "Solutrean", " ", " ", " "),
  Variable = c("Weight", "Length", "Width", "(n)","Weight", "Length", "Width","(n)", "Weight", "Length", "Width", "(n)"),
  L_median = c("1.4", "27.4", "10.7", "17", "1.2", "29", "11", "31", "1.25", "28", "11.4", "98"),
  NL_median = c("0.6", "23.7", "9.01", "9", "0.7", "21.5", "10.4", "13", "0.9", "28", "10.2", "15"),
  P_Value = round(c(res_ew_grav$p.value, res_el_grav$p.value, res_ewi_grav$p.value, 0, res_ew_proto$p.value, res_el_proto$p.value, res_ewi_proto$p.value, 0, res_ew_s$p.value, res_el_s$p.value, res_ewi_s$p.value, 0), 5)) |>   mutate(Significance = case_when(
      P_Value <= 0.001 ~ "***",
      P_Value <= 0.01  ~ "**",
      P_Value <= 0.05  ~ "*",
      TRUE             ~ ""),
    P_value = paste0(format(P_Value, nsmall = 4), Significance)  # Keep decimal places
  ) |> 
  select(-P_Value, -Significance) |> 
  dplyr::mutate(P_value = case_when(
    P_value == "0.00000***" ~ "-",
    TRUE ~ P_value))
